---
title: "FIT Bart to Curia 10x"
output: html_notebook
---

Split the curia datasets into practice and practice-year datasets so we can use the the same way.

```{r}
library(tidyverse)

knitr::opts_chunk$set(message = FALSE)

merged_dataset_practice_0026b <- read_csv("Data/curia_data/df_merged_raw_practice_level_10x/merged_dataset_practice_0026b.csv")
curia_merged_raw <- merged_dataset_practice_0026b

```

Are the columns unique? Return an empty logical if true, or a named list of the duplicates (second-occurring only of each pair).

```{r}
dup_cols <- duplicated(t(curia_merged_raw))

# Print out results
keep(dup_cols, isTRUE)

# Remove dupes from data
curia_merged_raw <- curia_merged_raw[!dup_cols]
```

Which cols vary at the practice level, and which at the practice year level? Use these variable naes to separate into `practice` and `practice_year` datasets

```{r}
# Get number of unique values per variable w/i practice. Practice level variables will be 1
practice_level_vars <- curia_merged_raw %>% 
  group_by(id.practice) %>% 
  summarise(
    n = n(),
    across(!n, n_distinct)
  ) %>% 
  select(-id.practice, -Z) %>% 
  {as.data.frame(t(rbind(
    mean = summarise(., across(everything(), mean)) %>% round(1),
    sd = summarise(., across(everything(), sd)) %>% round(3)
  )))} %>% 
  filter(mean == 1, sd == 0) %>% 
  rownames()

# Show results
practice_level_vars

# Create the two datasets
curia_practice <- curia_merged_raw %>% 
  select(id.practice, Z, all_of(practice_level_vars))
curia_practice_year <- curia_merged_raw %>% 
  select(!all_of(practice_level_vars))
```
# Functionalize

```{r}

```

