---
title: "R Notebook"
output: html_notebook
---

# Setup/metadata

```{r}
library(tidyverse)

knitr::opts_chunk$set(message = FALSE)

```

Which datasets are we looking at? I.e., which datasets did Curia provide us with?

```{r message=FALSE}
dataset_nums <- read_csv("Data/curia_data/dataset_nums.csv") %>% 
  pull(dataset_num)
```

Ground truth datasets

```{r}
estimand_truths <- read_csv("Data/ACIC_estimand_truths.csv")

```

DGP descriptions. Output of chunk shows unique combinations in current set of dataset nums.

```{r}
dgp_levels <- estimand_truths %>% 
  select(dataset.num:`Idiosyncrasy of Impacts`) %>% 
  distinct()

dgp_levels %>% 
  filter(dataset.num %in% dataset_nums) %>% 
  distinct(pick(-dataset.num))

```



## Practice/Practice-Year Information

```{r, eval=FALSE}
# Read in the competition data to get practice-year-level data for weighted means
n.patients.all <- dir("Data/track2_20220404/practice_year/", full.names = TRUE) %>% 
  map(~ read_csv(.x) %>% 
        select(id.practice, n.patients, year) %>% 
        mutate(dataset.num = str_match(basename(.x), '[0-9]+')[1,1])) %>% 
  list_rbind()

# Read in the competition data to get practice-level data for weighted means
practice_level <- dir("Data/track2_20220404/practice/", full.names = TRUE) %>% 
  map(~ read_csv(.x) %>% 
        mutate(dataset.num = str_match(basename(.x), '[0-9]+')[1,1])) %>% 
  list_rbind()

write_csv(practice_level, "Data/formatted/practice_true.csv")
write_csv(n.patients.all, "Data/formatted/n_patients_all.csv")

```

```{r}
# Load results of previous chunk
practice_level <- read_csv("Data/formatted/practice_true.csv")
n.patients.all <- read_csv("Data/formatted/n_patients_all.csv")

```

Subset number of patients to get weights

```{r}
# aggregate and summarize by all the variables
n.patients.overall <- n.patients.all %>% 
  group_by(dataset.num) %>% 
  summarize(n.patients = sum(n.patients), variable = "Overall")

n.patients.by_practice <- n.patients.all %>% 
  group_by(dataset.num, id.practice) %>% 
  summarise(n.patients = sum(n.patients))

n.patients.by_year <- n.patients.all %>% 
  group_by(dataset.num, year) %>% 
  summarise(n.patients = sum(n.patients), variable = "Overall")

# Merge on practice level variables
n.patients.practice_level <- n.patients.all %>% 
  left_join(practice_level, by = c('id.practice', 'dataset.num'))

# summarize by level of covariates
n.patients.by_X <- map(1:5, function(i) {
  n.patients.practice_level %>% 
    group_by(dataset.num, !!sym(paste0("X", i))) %>% 
    pivot_longer(cols = c(!!sym(paste0("X", i))), names_to = 'variable', values_to = 'level') %>% 
    mutate(level = as.character(level)) %>% 
    select(id.practice, n.patients, year, dataset.num, variable, level)
}) %>% 
  list_rbind() %>% 
  group_by(dataset.num, variable, level) %>% 
  summarise(n.patients = sum(n.patients))

# Create one dataframe to merge
n.patients.combined <- bind_rows(n.patients.overall, n.patients.by_practice, 
                                 n.patients.by_X, n.patients.by_year)
```


# Performance estimates

## ACIC winners

```{r, cache=TRUE}
acic_winners.raw <- readRDS("Data/eval_simple_tosend.RDS")
```

```{r}
acic_winners.df <- names(acic_winners) %>% 
  #keep(~ grepl("^t2", .x)) %>% 
  map(~ acic_winners[[.x]]$eval %>% mutate(method = .x)) %>%
  list_rbind()

```


## Curia

```{r, eval=FALSE}
# Load in curia estimates from separate tables.
# ⚠️ Slow
curia_estimates <- dir("Data/curia") %>%
  map(~ read_csv(file.path("Data", "curia", .))) %>%
  grep("acic_[a-z]+_([0-9]{4})\\.csv", ., value=T) %>% 
  list_rbind() %>%
  mutate(dataset.num = sprintf("%04d", dataset.num))

# Write for quick loading
write_csv(curia_estimates, "Data/all_curia_estimates.csv")

```

```{r}
# Load results of previous chunk
curia_estimates <- read_csv("Data/formatted/all_curia_estimates.csv")

```

### Performance metrics

```{r}
# Join on the ground truths
combined_ests <- curia_estimates %>% 
  left_join(estimand_truths, 
            by = c("dataset.num", "variable", "level", "year", "id.practice"),
            suffix = c(".curia", "")) %>% 
  select(!ends_with(".curia")) %>% 
  # Replace inner_join with left_join when running full scale on all dataset.num
  left_join(n.patients.combined, by = c("dataset.num", "variable", "level", "year", "id.practice")) %>% 
  filter(dataset.num %in% dataset_nums)

curia_eval.datasets <- combined_ests %>% 
  # Not estimating practice effects for this comparison
  filter(is.na(id.practice)) %>% 
  # Align names with scoring data from competition
  rename(true = SATT, satt = model_satt, se = model_se) %>% 
  mutate(bias = true - satt,
         abs_bias = abs(bias),
         # lower.90 = satt - qnorm(.95) * se,
         # upper.90 = satt + qnorm(.95) * se,
         # cover = between(true, lower.90, upper.90),
         width = qnorm(.95) * 2 * se) %>% 
  summarise(.by = c(variable, level, year, dataset.num),
            #rmse = caret::RMSE(satt, true),
            across(c(true, satt, se, bias:width), ~ weighted.mean(.x, n.patients))) %>% 
  mutate(lower.90 = satt - qnorm(.95) * se,
         upper.90 = satt + qnorm(.95) * se,
         cover = between(true, lower.90, upper.90)) %>% 
  mutate(id.practice = NA,
         method = "curia")
```


## Bart

```{r}
bart_satt_ests <- readRDS("Save/bart_satt_ests_600_1x.rds")

```

```{r}
bart_perf <- bart_satt_ests %>% 
  # Merge on the truth
  left_join(estimand_truths %>% 
              filter(is.na(year)), 
            by = c("dataset.num", "variable", "level", "id.practice"),
            suffix = c(".bart", "")) %>% 
  select(!ends_with(".bart")) %>% 
  rename(lower.90 = q05, upper.90 = q95, satt = est_satt, true = SATT) %>% 
  left_join(n.patients.combined, by = c("dataset.num", "variable", "level", "year", "id.practice")) %>% 
  filter(dataset.num %in% dataset_nums) %>% 
  mutate(bias = true - satt,
         abs_bias = abs(bias),
         # TODO: vvv why comment this line out? vvv
         # cover = between(true, lower.90, upper.90),
         width = upper.90 - lower.90) %>% 
  summarise(.by = c(variable, level, year, dataset.num, method),
            # TODO: vvv why comment this line out? vvv
            # rmse = caret::RMSE(satt, true),
            across(c(true, satt, bias:width, lower.90, upper.90), 
                   ~ weighted.mean(.x, n.patients))) %>% 
  # TODO: vvv Why calc this after summarize, instead of before? vvv
  mutate(cover = between(true, lower.90, upper.90)) %>% 
  mutate(id.practice = NA, se = NA)

  
```


# Merge the performance data

```{r}
# One large dataset for comparing all evals
evals_compare <- acic_winners_df %>% 
  select(-DGP_name) %>% 
  rbind(curia_eval.datasets) %>% 
  rbind(bart_perf) %>% 
  left_join(n.patients.combined, by = c("dataset.num", "variable", "id.practice", "level", "year")) %>% 
  select(-id.practice) %>% 
  filter(dataset.num %in% dataset_nums) %>% 
  # Merge back on DGP info
  left_join(dataset_dgp, by = 'dataset.num')

evals_compare

```


# Analysis and plots

```{r}
data.plot <- evals_compare %>% 
  # Summarize across datasets
  summarise(
    .by = c(variable, level, year, method, `Confounding Strength`),
    subgroup_size = mean(n.patients, na.rm = TRUE),
    rmse = caret::RMSE(satt, true),
    across(c(bias, abs_bias, cover, width), ~weighted.mean(., n.patients))
  ) %>% 
  filter(is.na(year)) %>% 
  # Remove unfinished ACIC submissions
  filter(!(method %in% c('t1_rBARTimpT', 'ofpsbart', 't1_H-TMLE'))) %>%
  # Create sensible variables for visualization
  mutate(
    # So we can treat all subgroups with the same variable
    variable_level = paste0(variable, 
                            ifelse(is.na(level), '', paste0('_', level)),
                            ifelse(is.na(year), '', paste0('_y', year))),
    # Ensures methods of interest are at the top of the draw order
    method = factor(method, levels = rev(c('curia', 'BART', 'DART', 
                                           setdiff(unique(method), 
                                                   c('curia', 'BART', 'DART'))))),
    # Sensible method labelling
    method_label = case_match(method, 
                              'curia' ~ 'curia', 
                              'BART' ~ 'BART',
                              'DART' ~ 'DART',
                              .default = 'other')
  )

data.plot

method_colors <- c(Curia = "#F8766D", Oracle = "#00BFC4", Other = "darkgrey", 
                   BART = '#4356FF', DART = '#AB54F9')

```

```{r}
ggplot(data = data.plot, aes(x = subgroup_size, y = rmse)) + 
         geom_line(aes(group = method, color = method_label)) +
  scale_color_manual(values = c(curia = 'darkblue', BART = 'darkred', DART = 'cyan', other = 'lightgray')) +
  facet_wrap(~ `Confounding Strength`) +
  scale_y_log10()
```

```{r, eval=FALSE}
evals_compare %>% 
         summarise(.by = c(variable, level, year, method),
                   subgroup_size = mean(n.patients),
                   rmse = caret::RMSE(satt, true), 
                   across(c(satt, true, bias, abs_bias, lower.90, upper.90, cover, width),
                          ~ weighted.mean(., n.patients))) %>%
        write_csv("Data/eval_compare.csv")
```

```{r}
#method_groups = c('curia', keep(names(acic_winners), ~grepl("^t2", .)))

map(c("rmse", "bias", "abs_bias", "cover", "width"), function(.x) {
  ggplot(data = data.plot, aes(x = subgroup_size, y = !!sym(.x))) + 
    geom_line(aes(group = method, color = method_label)) +
    scale_color_manual(values = c(curia = 'darkblue', BART = 'darkred', DART = 'cyan', other = 'lightgray')) +
    facet_wrap(~ `Confounding Strength`) +
    {if(.x %in% c("rmse")) scale_y_log10() else NULL} + 
    labs(title = .x)
  ggsave(file = paste0("Plots/plot_", .x, "_by_subgroup_size.png"))
})

```

```{r}
# Mare's plot_perf code
perf <- data.plot %>%
  filter(!(method %in% c('t1_rBARTimpT', 'ofpsbart', 't1_H-TMLE'))) %>%
  filter(variable=='Overall' & is.na(year))

# perf <- rbind(perf, rep(NA, ncol(perf)))
# perf$method[nrow(perf)] <- 'Oracle'
# perf$rmse[nrow(perf)] <- 0
# perf$width[nrow(perf)] <- 0
# perf$cover[nrow(perf)] <- .9

oracle_row <- data.frame(
  method = "Oracle",
  rmse = 0,
  width = 0,
  cover = 0.9,
  variable = "Overall",
  bias = 0,
  abs_bias = 0
)
perf <- bind_rows(perf, oracle_row) %>% 
  mutate(Method = case_match(
    method,
    'curia' ~ 'Curia',
    'Oracle' ~ 'Oracle',
    'BART' ~ 'BART',
    'DART' ~ 'DART',
    .default = "Other"
  ))

# perf <- perf %>%
#   mutate(Method = ifelse(method == 'curia', 'Curia', 
#                          ifelse(method == 'Oracle', 'Oracle', 'Other')))

# method_colors <- c(Curia = "#F8766D", Oracle = "#00BFC4", Other = "darkgrey")
method_colors <- c(Curia = "#F8766D", Oracle = "#00BFC4", Other = "darkgrey", 
                   BART = '#4356FF', DART = '#AB54F9')


#meta <- readRDS('real_submissions.RDS')

curia_rmse <- perf %>%
  filter(Method=='Curia') %>%
  pull(rmse)

perf %>%
  summarise(percent_better = mean(rmse < curia_rmse, na.rm=T))

perf %>%
  ggplot(aes(x=rmse, y=cover, color=Method)) + 
  geom_hline(yintercept = 0.9, color='darkgrey') +
  geom_point() + 
  xlab('RMSE, overall SATT') +
  ylab('Coverage of overall SATT 90% interval') +
  facet_wrap(~ `Confounding Strength`) +
  scale_color_manual(values=method_colors)
ggsave(file.path('Plots', 'rmse_cover.png'), height=3.5, width=4.5)

perf %>%
  ggplot(aes(x=width, y=cover, color=Method)) + 
  geom_hline(yintercept = 0.9, color='darkgrey')+
  geom_point() + 
  xlab('Width of overall SATT 90% interval') +
  ylab('Coverage of overall SATT 90% interval') +
  facet_wrap(~ `Confounding Strength`) +
  scale_color_manual(values=method_colors)
ggsave(file.path("Plots", 'width_cover.png'), height=3.5, width=4.5)

```


